---
title: "2025S2 CI-GroupAssignment-report"
author: "ML Abrahams"
date: "2025-03-17"
output: pdf_document
---
```{r}
#setwd("C:\\Users\\leoni\\OneDrive\\Bureaublad\\ACADEMIA\\Statistics & Data Science MSc\\CAUSAL INFERENCE 1\\Group assignment")
data = read.csv("2025_framingham_assignment.csv")
library(ggplot2)
library(car)
```
#1: Data preperation
ADJUST DATASET TO SELECTED VARIABLES BY DAG
```{r data-preparation}
# Set SEX to binary variable with 0 (male) and 1 (female)
data$SEX = ifelse(data$SEX == 1, 0, 1)

# Get columns specific for answering our research questions
data = data[,c(2:3,5:11,14,16)]

# Identify NA's
getNA = function(data) {
  sum(is.na(data))
}

NA.total = apply(data, 2, getNA)
NA.total # 113 NA in EDUC variable

# Exclude entries where EDUC is NA
removeNA = function(data) {
  if(is.na(data[8]) == FALSE) {
    keep = data
  }
}

data.noNA = data[!is.na(data$EDUC),]

# Check again for NA distribution
NA.total.1 = apply(data.noNA, 2, getNA)
NA.total.1
```
#2: Fit outcome regression model
ADJUST MODEL BASED ON THE RIGHT VARIABLES
```{r fit-outcome-regression}
# Fit outcome regression models on data set with all variables including interactions between all variables and EDUC
# AND na.exclude
outcome.model = glm(STROKE ~
                      SEX * EDUC +
                      AGE * EDUC +
                      DIABP * EDUC +
                      BPMEDS * EDUC +
                      CURSMOKE * EDUC +
                      DIABETES * EDUC +
                      BMI * EDUC +
                      HEARTRTE * EDUC +
                      HYPERTEN * EDUC, 
                    data = data.noNA, 
                    family = binomial, 
                    na.action = na.exclude)

summary(outcome.model)
```
# 3: Find ATE
```{r find-ATE}
# Duplicate data sets to two sets with different exposure levels and predict individual E(Y = y)
## Set all to exposure 1 (EDUC == 1)
data.1 = data.noNA
data.1$EDUC = 1 # high education

## Set data set to exposure 0 (EDUC == 0)
data.0 = data.noNA
data.0$EDUC = 0 # low education

# Get predicted values for E(Y = y | X = x) based on outcome regression model
getPred = function(dataset) {
  pred = predict(outcome.model, dataset, type = "response")
  out = mean(pred, na.rm = TRUE)
  return(out)
}

# Predict values E(Y = y | X = 1) based on outcome regression model
EY1 = getPred(data.1)

# Predict values E(Y = y | X = 0) based on outcome regression model
EY0 = getPred(data.0)

# Find average treatment effect (effect of Education on stroke)
ATE = EY1 - EY0 #ATE = -.01 > people with a higher education have 1% less change of developing a stroke
c(ATE, EY1, EY0)

# Standardization using stdGlm
library(stdReg)
fit.std = stdGlm(fit = outcome.model, data = data.noNA, X = "EDUC")
summary(fit.std, contrast = "difference", reference = 0)
```
#4: Assumptions
##4.1: For CI

##4.2: Outcome regression model
For logistic regression, there is no assumption on the error terms. The assumtptions that are made though are:

* Linearity of the log-odds (independent and log-odds are linear)
* Independent observations (assumed by the dataset, no further testing)
* No perfect multicollinearity
* Correct specification

The error terms do show a specific patterns, which is additionally clustered per outcome value. This indicates a misspecification of the model (missing variable or interaction). The model does not depend on these error terms, but there might be something wrong with the model. 
```{r testing-assumptions}
# Conditional exchangeablility

# Positivity

# Consistency

# OUTCOME REGRESSION MODEL SPECIFIC
## Linearity in log odds (dependent and independent variables)
##

## High influential points
vif(outcome.model)

## Homoscedasticity of error terms 
fitted_vals = fitted(outcome.model)    # Extract fitted values
residuals_vals = residuals(outcome.model)  # Extract residuals

plot(fitted_vals, residuals_vals, 
     col = ifelse(data.noNA$STROKE == 1, "blue", "red"),
     pch = 16, 
     xlab = "Fitted Values", 
     ylab = "Residuals",
     main = "Residual Plot Colored by Outcome (0 = Red, 1 = Blue)")
```
